Running Anpan

### 2023-03-30 other/anpan.qmd

@@

After identifying samples with *B. wexlerae* in our sample set,
I generated a table with stratified gene families by running

```fish
echo -e "sample\tfeature\tabundance" > data/bwexlerae_genefamilies.tsv
for s in (cat has_bwex.txt)
    fd "$s"_genefamilies /grace/sequencing/processed/mgx/humann/main/ -x rg s__Blautia_wexlerae | sd '^q' "$s\t" >> data/bwexlerae_genefamilies.tsv
end

sd '^.+\t(.+\t.+\t.+)$' '$1' data/bwexlerae_genefamilies.tsv
sd '^FG\d+_S\d+\t$' '' data/bwexlerae_genefamilies.tsv

and then added a header manually.

```{r}
library(data.table)

df <- fread("data/bwexlerae_genefamilies.tsv")
df2 <- dcast(df[, .(feature, sample, abundance)],
                 feature ~ sample,
                 value.var = "abundance",
                 fun.aggregate = sum)
names(df2)[1] = "# Gene Family"
```

@@

```{r}
bwex_results <- anpan(#bug_dir           = "/grace/sequencing/processed/mgx/humann/main",
                        bug_file          = "./data/bwexlerae_genefamilies_wide.tsv",
                        meta_file         = "./data/manuscript_meta.csv",
                        out_dir           = "./data/anpan",
                        # annotation_file   = "/path/to/annotation.tsv", #optional, used for plots
                        filtering_method  = "kmeans",
                        model_type        = "fastglm",
                        covariates        = c("ageMonths", "sex", "education"),
                        outcome           = "cogScore",
                        plot_ext          = "pdf",
                        save_filter_stats = TRUE)
```

```{r}
input_path = file.path("data", "anpan", "filter_stats",
                       "filtered_bwexlerae_genefamilies_wide.tsv")

@@

>>>>

## Identifying taxa

```{julia}
using Resonance

species = Resonance.load(TaxonomicProfiles())

bugs = [
    "Blautia_wexlerae",
    "Ruminococcus_gnavus",
    "Erysipelatoclostridium_ramosum",
    "Bifidobacterium_longum"
]

for bug in bugs
    ss = filter(s-> species[Taxon(bug, :species), s] > 0, samples(species))
    open(joinpath("data", "anpan", "has_" * lowercase(replace(bug, r"([A-Z])[a-z]+_([a-z]{3}).+"=> s"\1\2")) * ".txt"), "w") do io
        println.(io, name.(ss))
    end
end
```

```{julia}
mdata = Resonance.load(Metadata())
pathways = Resonance.load_raw_humann(; kind="pathabundance", sample_filter=mdata.sample, stratified = true)

for bug in bugs
    comm = pathways[findall(f-> !ismissing(taxon(f)) && replace(name(taxon(f)), r".+s__(.+)"=> s"\1") == bug, features(pathways)), :]
    df = DataFrame(comm)
    df.features = string.(df.features)
    rename!(df, "features"=> "# Gene Family")
    df = df[!, findall(col-> col == "# Gene Family" || any(>(0), df[!, col]), names(df))]
    CSV.write(joinpath("data", "anpan", bug * "_pathabundance.tsv"))
end
```

@@

After identifying samples with *B. wexlerae* in our sample set

```{r}
library(data.table)
```

@@

## Blautia_wexlerae

```{r}
bwex_results <- anpan(#bug_dir           = "/grace/sequencing/processed/mgx/humann/main",
                        bug_file          = "./data/anpan/Blautia_wexlerae_pathabundance.tsv",
                        meta_file         = "./data/anpan/manuscript_meta.csv",
                        out_dir           = "./data/anpan/Blautia_wexlerae",
                        # annotation_file   = "/path/to/annotation.tsv", #optional, used for plots
                        filtering_method  = "kmeans",
                        model_type        = "fastglm",
                        covariates        = c("ageMonths", "sex", "education"),
                        outcome           = "cogScore",
                        plot_ext          = "pdf",
                        save_filter_stats = TRUE)
```

```{r}
input_path = file.path("data", "anpan", "Blautia_wexlerae", "filter_stats",
                       "filtered_Blautia_wexlerae_pathabundance.tsv.gz")

@@

## Ruminococcus_gnavus

```{r}
results <- anpan(#bug_dir           = "/grace/sequencing/processed/mgx/humann/main",
                        bug_file          = "./data/anpan/Ruminococcus_gnavus_pathabundance.tsv",
                        meta_file         = "./data/anpan/manuscript_meta.csv",
                        out_dir           = "./data/anpan/Ruminococcus_gnavus",
                        # annotation_file   = "/path/to/annotation.tsv", #optional, used for plots
                        filtering_method  = "kmeans",
                        model_type        = "fastglm",
                        covariates        = c("ageMonths", "sex", "education"),
                        outcome           = "cogScore",
                        plot_ext          = "pdf",
                        save_filter_stats = TRUE)

input_path = file.path("data", "anpan", "Ruminococcus_gnavus", "filter_stats",
                       "filtered_Ruminococcus_gnavus_pathabundance.tsv.gz")

plot_results(res         = results,
             model_input = model_input,
             covariates  = c("ageMonths", "sex", "education"), 
             outcome     = "cogScore",
             bug_name    = "R. gnavus",
             cluster     = "both",
             show_trees  = TRUE)

## Erysipelatoclostridium_ramosum

```{r}
results <- anpan(#bug_dir           = "/grace/sequencing/processed/mgx/humann/main",
                        bug_file          = "./data/anpan/Erysipelatoclostridium_ramosum_pathabundance.tsv",
                        meta_file         = "./data/anpan/manuscript_meta.csv",
                        out_dir           = "./data/anpan/Erysipelatoclostridium_ramosum",
                        # annotation_file   = "/path/to/annotation.tsv", #optional, used for plots
                        filtering_method  = "kmeans",
                        model_type        = "fastglm",
                        covariates        = c("ageMonths", "sex", "education"),
                        outcome           = "cogScore",
                        plot_ext          = "pdf",
                        save_filter_stats = TRUE)

input_path = file.path("data", "anpan", "Erysipelatoclostridium_ramosum", "filter_stats",
                       "filtered_Erysipelatoclostridium_ramosum_pathabundance.tsv.gz")

plot_results(res         = results,
             model_input = model_input,
             covariates  = c("ageMonths", "sex", "education"), 
             outcome     = "cogScore",
             bug_name    = "E. ramosum",
             cluster     = "both",
             show_trees  = TRUE)

## Bifidobacterium_longum

```{r}
results <- anpan(#bug_dir           = "/grace/sequencing/processed/mgx/humann/main",
                        bug_file          = "./data/anpan/Bifidobacterium_longum_pathabundance.tsv",
                        meta_file         = "./data/anpan/manuscript_meta.csv",
                        out_dir           = "./data/anpan/Bifidobacterium_longum",
                        # annotation_file   = "/path/to/annotation.tsv", #optional, used for plots
                        filtering_method  = "kmeans",
                        model_type        = "fastglm",
                        covariates        = c("ageMonths", "sex", "education"),
                        outcome           = "cogScore",
                        plot_ext          = "pdf",
                        save_filter_stats = TRUE)

input_path = file.path("data", "anpan", "Bifidobacterium_longum", "filter_stats",
                       "filtered_Bifidobacterium_longum_pathabundance.tsv.gz")

plot_results(res         = results,
             model_input = model_input,
             covariates  = c("ageMonths", "sex", "education"), 
             outcome     = "cogScore",
             bug_name    = "B. longum",
             cluster     = "both",
             show_trees  = TRUE)
